---
title: "How to Use Citation Manager"
subtitle: "Featuring Mendeley & Zotero"
author: "Kadek Adit Wiryadana"
institute: "Intisari Sains Medis"
date: "2021/04/04 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      titleSlideClass: [center, middle]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# Citation Manager?

Reference is an **integral** part of any Scientific Paper.
However, creating one, *a good and a right reference*, is tiring and laborious.

**Here comes the CITATION MANAGER**
* Ease the burden to create a consistent reference
* Automatic update on the inline citation
* Save metadata online and offline
* Importing metadata online, one click away

---
# Available Solution

There are plenty of software we could use. Here are some of them:

* **Endnote** -> Paid and Proprietary Software (by Clarivate Analytics)

* **Mendeley** -> Free but Proprietary Software (by Elsevier)

* **Zotero** -> Free and Open-source Software (by Digital Scholarship Foundation)

.footnote[[*] We will talk about Mendeley and Zotero]

---
# Set of Software Environment

There are basically three set of main apps

1. Desktop Client (offline apps)

2. Online repository (like the *google drive* of citation metadata)

3. Browser add-on (the bridge between offline-online apps, and act as web importer)

Also, each of software solution provide its mobile apps (Android and IOS) to access the online repository.

---
# Installation: Desktop Client
```{r installation, echo=FALSE}
xaringanExtra::use_panelset()
```
.panelset[
.panel[.panel-name[Windows]

1. Register Account on [Mendeley Website](https://www.mendeley.com/)

2. Download Installer on the [website](https://www.mendeley.com/download-desktop-new/)

3. Install the program by double clicking the installer

4. Follow the usual installation promt
]

.panel[.panel-name[Linux]

1. Open Terminal

2. type "sudo apt-get install mendeleydesktop"

3. Done
]

.panel[.panel-name[MacOs]

1. I dont have MacOs

2. So I think it is similar to Windows

3. Please be so.....

]
]
---
# Installation of Web Importer: Firefox
Web importer is a piece of apps that run on the browser act as a bridge to pass metadata from website to your online repository.
```{r web importer, echo=FALSE}
xaringanExtra::use_panelset()
```
.panelset[
.panel[.panel-name[Firefox]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/rFmHr9YHmrw" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]
.panel[.panel-name[Chrome]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/NcKxDw_ik7U" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]  
]

---
# CSL Template
Mendeley and other citation manager use CSL (citation style languange) as the template of citation and bibliographics format.
I have prepared a custom CSL for Intisari Sains Medis, stored on Mendeley server

"https://www.mendeley.com/reference-management/web-importer"

Here is the guide of how to add this CSL to your Mendeley style collection!

.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/S00p5Socf_w" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
]
---
# Adding Metadata
Mendeley used the metadata (identity of files) to compose the bibliography.

There are basically three ways of adding metadata. Take a look of these videos.
```{r add-metadata, echo=FALSE}
xaringanExtra::use_panelset()
```
.panelset[
.panel[.panel-name[Web-Importer]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/T_JK6H4g2l4" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]
.panel[.panel-name[Manual]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/STejkQE92E0" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]  
]

---
# Main Course: Inline Citation and Bibliography
Finally,  adding citation to our working document. There are 2 steps:
1. First click where would you like to put the inlice citation in your working document, then click add citation/reference in the mendeley's toolbar.
2. Add bibliography in the reference section of your working document. 
```{r add-citation, echo=FALSE}
xaringanExtra::use_panelset()
```
.panelset[
.panel[.panel-name[Libre-Office]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/5hBLYa3LJF0" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]
.panel[.panel-name[Microsoft-Office]
.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/v_1-om0s9l4" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
    ]
  ]  
]

---
# Correction and Automatic Update
The most powerful advantage of citation manager is the automatic update of your inline citation and bibliography when the documents edited or the metadata changed.
Add once, and then you can focus on the content.

.center[
<iframe width="560" height="315" src="https://www.youtube.com/embed/nQiNwSCR1zU" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
]
---
# Few Examples
```{r examples, echo=FALSE}
xaringanExtra::use_panelset()
```
.panelset[
.panel[.panel-name[Inline Citation]
.left[
.center[** This is the inline citation**]

First sentence cited from a single document.<sup>1</sup>
Then the second sentence were based on two other documents.<sup>2,3</sup>
We can combine these documents which support your hypothesis.<sup>1-3</sup>
Of course, there are other documents that doesnt support your hypothesis.<sup>4-9</sup>
Finally, the closing sentence were summaries of all the relevant fact except from documents 2.<sup>1,3-9</sup>
    ]
  ]

.panel[.panel-name[Bibliography]

.center[** This is the bibliograpy in Reference Section**]
.left[
Journal Article

1. Tini K, Samatra IDPGP, Wiryadana KA, Supadmanaba IGP. Clinical Profile of Patients with Cerebrovascular Disease at Stroke Center, Sanglah Hospital Denpasar Bali. *Bali Med J*.2020;9(1):129–36. Available from: https://www.balimedicaljournal.org/index.php/bmj/article/view/1665

Books

2. Jameson J, Fauci AS, Kasper DL, Hauser SL, Longo D, Loscalzo J. Harrison’s Principles of Internal Medicine, 20th ed. McGraw-Hill. New York: Mcgraw-Hill Education; 2018.

Books Section

3. Crainiceanu C, Caffo B, Muschelli J. Hyphothesis Testing. In: Methods in Biostatistics with R. British Columbia: Leanpub; 2020. p. 487–506.
    ]
  ]
  
.panel[.panel-name[Bibliography-2]
.center[** This is the bibliograpy in Reference Section**]
.left[
Computer Program

4. R Core Team. R: A language and environment for statistical computing. [Internet]. Vienna: R Foundation for Statistical Computing; 2020. Available from: https://www.r-project.org/

Web Page

5. Grolemund G, Wickham H. R for Data Science [Internet]. Available from:
https://r4ds.had.co.nz/

Conference Proceeding

6. Scepanovic S, Martin-Lopez E, Quercia D, Baykaner K. Extracting medical entities from
social media. In: ACM CHIL 2020 - Proceedings of the 2020 ACM Conference on Health,
Inference, and Learning. Association for Computing Machinery, Inc; 2020. p. 170–81.

    ]
  ] 
.panel[.panel-name[Bibliography-3]
.center[** This is the bibliograpy in Reference Section**]
.left[ 

Report

7. WHO/Department of Control of neglected tropical diseases. Investing to overcome the
global impact of neglected tropical diseases [Internet]. Geneva; 2015. Available from:
https://www.who.int/neglected_diseases/9789241564861/en/

Thesis

8. Supadmanaba IGP. The effect of Anemia to Cardiovascular Risk, Lipid Profile, and
Glucose Monitoring in Patient with type 2 Diabetes Mellitus in Sanglah General Hospital.
Universitas Udayana; 2014.
    ]
  ]  
]

---
# Links/Source

The CSL files can be download my github repository, link [here](https://github.com/wiryadana/Xaringan_citation/blob/main/vancouver-superscript-IntisariSainsMedis.csl) or from [mendeley server](https://csl.mendeley.com/styles/315880411/vancouver-superscript-IntisariSainsMedis)

The full documentation, written in Bahasa Indonesia, could be accessed from [here](https://drive.google.com/file/d/1qjfj0BVxv1c9MXmPRzsa-Y2jQYT-IO29/view?usp=sharing)*

---
class: center, middle

# Thanks!
full document can be accesed on --- 

**Any Question or Suggestion are Welcomed**

Kadekaditwiryadana@hotmail.com or 
fbihansip@gmail.com

.footnote[Slides created with R and R package [**xaringan**](https://github.com/yihui/xaringan) in Rstudio.]

